---
title: "p8105_hw5_yh3964"
author: "Yirou Hu"
date: "2025-11-02"
output: github_document
---
```{r}
library(tidyverse)
theme_set(theme_minimal() + theme(legend.position = "bottom"))  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1
Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

- write a function that, for a fixed group size, randomly draws “birthdays” for each person; 
- checks whether there are duplicate birthdays in the group; 
- returns TRUE or FALSE based on the result.

```{r}
# Define simulation birthday
birthday_sim = function(n_room) {
  birthdays = sample(1:365, n_room, replace = TRUE)
  rep_birthday = length(unique(birthdays)) < n_room
  rep_birthday
}
```

