---
title: "p8105_hw5_yh3964"
author: "Yirou Hu"
date: "2025-11-02"
output: github_document
---
```{r}
library(tidyverse)
theme_set(theme_minimal() + theme(legend.position = "bottom"))  
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1
Suppose you put n people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

- write a function that, for a fixed group size, randomly draws “birthdays” for each person; 
- checks whether there are duplicate birthdays in the group; 
- returns TRUE or FALSE based on the result.

## Define function
```{r}
birthday_sim = function(n_room) {
  birthdays = sample(1:365, n_room, replace = TRUE)
  rep_birthday = length(unique(birthdays)) < n_room
  rep_birthday
}
```

- Run this function 10000 times for each group size between 2 and 50. 
- For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs. 
- Make a plot showing the probability as a function of group size, and comment on your results

## simulation
```{r}
group_sizes = 2:50
n_simulations = 10000

birthday_prob = 
  tibble(group_size = group_sizes) |> 
  mutate(
    simulation_results = map(group_size, ~replicate(n_simulations, birthday_sim(.x))),
    prob_duplicate = map_dbl(simulation_results, mean)
  ) |> 
  select(group_size, prob_duplicate)
```

## Visualization
```{r}
birthday_plot = 
  birthday_prob |> 
  ggplot(aes(x = group_size, y = prob_duplicate)) +
  geom_line(linewidth = 1.2, color = "blue") +  
  geom_point(size = 2, color = "orange") +       
  labs(
    x = "Group Size (Number of People)",
    y = "Probability of At Least Two People Sharing a Birthday",
    title = "Relationship Between Group Size and Duplicate Birthday Probability",
  ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +  
  theme(plot.title = element_text(hjust = 0.5, size = 10))

print(birthday_plot)
```

## Comments

The plot vividly illustrates the birthday paradox phenomenon. 

- As the group size grows, the probability of at least two people sharing a birthday rises quickly. For example, when the group size reaches 23, the probability exceeds 50% for most people, who often expect a much larger group size to reach this threshold.

- When the group size approaches 50, the probability nears 100%, meaning it is almost guaranteed that at least two people will share a birthday in such a group.

- This result highlights how statistical probability can diverge from everyday intuition. The rapid rise is driven by the increasing number of pairwise comparisons.

## Problem 2

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

- Fix n = 30
- Fix $\sigma$ = 5
- Set $\mu$ = 0 
- Generate 5000 datasets from the model: $x ∼ normal [\mu, \sigma]$
- For each dataset, save $\hat{\mu}$ and the p-value arising from a test of H: $\mu$ = 0 using $\alpha$ = 0.05

```{r}
n_simulations = 5000
n = 30
sigma = 5
mu = 0
alpha = 0.05

# Define the simulation function
simulate_t_test = function(mu, n, sigma) {
  sample = rnorm(n, mean = mu, sd = sigma)
  test_result = t.test(sample, mu = 0, conf.level = 0.95)
  tidy_result = broom::tidy(test_result) |>
    rename(mu_hat = estimate, p_value = p.value)
  return(tidy_result[c("mu_hat", "p_value")])
}

# simulate for 5000 times for mu = 0
simulation_results = map_dfr(1:n_simulations, ~simulate_t_test(mu, n, sigma))
```

Repeat the above for $\mu$ ={1,2,3,4,5,6}, and complete the following:

- Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of $\hat{\mu}$ on the x axis. Describe the association between effect size and power.

- Make a plot showing the average estimate of $\hat{\mu}$ on the y axis and the true value of $\mu$ on the x axis. 

- Make a second plot (or overlay on the first) the average estimate of $\hat{\mu}$ only in samples for which the null was rejected on the y axis and the true value of $\mu$ on the x axis. Is the sample average of $\hat{\mu}$ across tests for which the null is rejected approximately equal to the true value of $\mu$ ? Why or why not?

```{r}
true_mu_values = c(1, 2, 3, 4, 5, 6)

all_sim_results = map_dfr(
  true_mu_values,  
  function(mu) {
    map_dfr(1:n_simulations, ~simulate_t_test(mu, n, sigma)) |>
      mutate(true_mu = mu)
  }
)

power_df = all_sim_results |>
  group_by(true_mu) |>
  summarize(
    power = mean(p_value < alpha),  
    .groups = "drop"
  )

# Power plot
power_plot = power_df |>
  ggplot(aes(x = true_mu, y = power)) +
  geom_point(size = 3, color = "darkblue") +
  geom_line(linewidth = 1.2, color = "darkblue") +
  labs(
    x = "True Population Mean (μ)",
    y = "Power (Probability of Rejecting Null)",
    title = "Power of One-Sample t-Test vs. Effect Size"
  ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  theme_minimal()

print(power_plot)


mu_hat_analysis = all_sim_results |>
  group_by(true_mu) |>
  summarize(
    avg_mu_hat_all = mean(mu_hat),  
    avg_mu_hat_rejected = mean(mu_hat[p_value < alpha], na.rm = TRUE),
    .groups = "drop"
  )

# Plot for average estimate of mu hat on the y axis and the true value of mu hat on the x axis
mu_hat_all_plot = mu_hat_analysis |>
  ggplot(aes(x = true_mu, y = avg_mu_hat_all)) +
  geom_point(size = 3, color = "blue") +
  geom_line(linewidth = 1.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  labs(
    x = "True Population Mean (μ)",
    y = "Average Estimate of μ hat",
    title = "Average μ hat vs. True μ (All Simulations)"
  ) +
  theme_minimal()

print(mu_hat_all_plot)

# Plot for all simulations vs rejected nulls
mu_hat_plot = mu_hat_analysis |>
  ggplot() +
  geom_point(aes(x = true_mu, y = avg_mu_hat_all), color = "blue", size = 3) +
  geom_line(aes(x = true_mu, y = avg_mu_hat_all), color = "blue", linewidth = 1.2) +
  geom_point(aes(x = true_mu, y = avg_mu_hat_rejected), color = "red", size = 3) +
  geom_line(aes(x = true_mu, y = avg_mu_hat_rejected), color = "red", linewidth = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  labs(
    x = "True Population Mean (μ)",
    y = "Average Estimate (μ hat)",
    title = "All Simulations μ hat vs. Rejected Nulls"
  ) +
  theme_minimal()

print(mu_hat_plot)
```

## Comments on three plots and answer questions

- Power vs. True $\mu$: The first plot illustrates a strong positive association between the true effect size $\mu$ and the power of the one-sample t-test. As $\mu$ increases (from 1 to 6), the probability of correctly rejecting the false null hypothesis rises sharply. Power plateaus at nearly 100% when $\mu \geq 4$, meaning further increases in $\mu$ do not meaningfully improve the ability to detect the effect.

- Average $\hat{\mu}$ Across All Simulations: The second plot (average for all tests) shows that the sample mean $\hat{\mu}$ is an unbiased estimator of the true $\mu$. The points lie almost exactly on the dashed 45° line, indicating that, on average, $\hat{\mu}$ equals the true $\mu$ across all simulated samples.

- Average $\hat{\mu}$ for Rejected Tests vs. All Tests: The third plot shows that the sample average of $\hat{\mu}$ for tests where the null hypothesis is rejected is consistently higher than the true $\mu$, particularly for smaller effect sizes. This implies that when the null hypothesis is rejected, the estimate of $\mu$ is biased upward. The bias arises from a selection effect: only samples with larger $\hat{\mu}$ values (due to random variation) result in the null being rejected. Thus, the sample average of $\hat{\mu}$ for rejected tests is not approximately equal to the true $\mu$, indicating an overestimation of the true effect in these cases.

# Problem 3
Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r}
homicide_data = read_csv("homicide-data.csv")

homicides_summary_df =  
  homicide_data |>
  janitor::clean_names() |>
  drop_na() |>
  mutate(city_state = paste(city, state, sep = ", ")) |>
  group_by(city_state) |>
  summarise(
    total_num_homicides = n(), 
    total_num_unsolved_homicides = sum(disposition == "Closed without arrest" | disposition == "Open/No arrest")) |> 
  filter(total_num_unsolved_homicides != 0)

homicides_summary_df
```
- The raw data includes `r nrow(homicide_data)` observations and `r length(names(homicide_data))` variables. 

- We have variables: uid, reported_date, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, lat, lon, and disposition.

- Each row represents a single homicide case occurring in one of 50 cities spanning 28 states.


- For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.
```{r}
baltimore_df =
  homicides_summary_df |>
  filter(city_state == "Baltimore, MD") 

baltimore_total = pull(baltimore_df, total_num_homicides)
baltimore_unsolved = pull(baltimore_df, total_num_unsolved_homicides)


baltimore_test_prop = prop.test(baltimore_unsolved, baltimore_total)

baltimore_summary = broom::tidy(baltimore_test_prop) |>
  mutate(
    estimate_p = estimate,
    confidence_low = conf.low,
    confidence_high = conf.high
  ) |>
  select(estimate_p, confidence_low, confidence_high)

baltimore_summary |> 
  knitr::kable()
```

- Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.
```{r}
all_city_estimates = homicides_summary_df |>
  group_by(city_state) |>
  summarize(
    prop_test = list(prop.test(total_num_unsolved_homicides, total_num_homicides))
  ) |>
  mutate(tidy_test = purrr::map(prop_test, broom::tidy)) |>
  unnest(tidy_test) |>
  mutate(
    estimate_p = estimate,
    confidence_low = conf.low,
    confidence_high = conf.high
  ) |>
  select(city_state, estimate_p, confidence_low, confidence_high)

all_city_estimates |> 
  knitr::kable()
```

- Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides
```{r}
ggplot(all_city_estimates, aes(x = reorder(city_state, estimate_p), y = estimate_p)) +
  geom_point(color = "blue", size = 1.5) +
  geom_errorbar(aes(ymin = confidence_low, ymax = confidence_high)) +
  coord_flip() +
  labs(
    x = "City", 
    y = "Proportion of Total Unsolved Homicides",
    title = "Estimated Unsolved Homicide Proportion by City",
    subtitle = "95% Confidence Intervals for All Cities"
    ) +
  theme(axis.text.y = element_text(size = 6))
```
The plot displays the estimated unsolved homicide proportion across U.S. cities, sorted by clearance rate (lowest to highest from bottom to top). 

Confidence intervals show precise estimates (narrow bars) for high-volume cities.

